Commutative Character Sheaves

### 2015-07-23 followup.tex
New file for followup paper.

\documentclass[10pt]{amsart}
\usepackage[a4paper]{geometry}
\title{From the function-sheaf dictionary to quasicharacters of $p$-adic tori}

\author{Clifton Cunningham}
\author{David Roe}

\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, Alberta, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\address{Department of Mathematics, University of British Columbia, 1984 Mathematics Road, Vancouver, British Columbia, Canada, {V6T~1Z2}.}
\email{roed.math@gmail.com}

\subjclass[2010]{14F05 (primary), 14L15 (secondary), 22E50 (tertiary)}

\keywords{character sheaves, p-adic tori, N\'eron models, Greenberg functor, geometrization, quasicharacter sheaves}

%%%%%%%%%%%%%%%%% PACKAGES %%%%%%%%%%%%%%%%%

\usepackage{amssymb}
\usepackage[initials]{amsrefs}

%Customizations for amsrefs
\BibSpec{lecture}{%
  +{} {\PrintAuthors}				{author}
  +{,}{ }						{title}
  +{}{ \PrintDateB}				{date}
  +{.}{ Lecture delivered at \textit}	{conference}
  +{,}{ }						{address}
  +{.}{ }						{transition}
}

% We make volume bold rather than vol. #
\BibSpec{book}{%
    +{}  {\PrintPrimary}				{transition}
    +{,} { }							{title}
    +{.} { }							{part}
    +{:} { }							{subtitle}
    +{,} { \PrintEdition}					{edition}
    +{}  { \PrintEditorsB}				{editor}
    +{,} { \PrintTranslatorsC}			{translator}
    +{,} { \PrintContributions}			{contribution}
    +{,} { }							{series}
    +{} { \textbf}						{volume}
    +{,} { }							{publisher}
    +{,} { } 							{organization}
    +{,} { }							{address}
    +{,} { \PrintDateB}					{date}
    +{,} { }							{status}
    +{}  { \parenthesize}				{language}
    +{}  { \PrintTranslation}				{translation}
    +{;} { \PrintReprint}				{reprint}
    +{.} { }							{note}
    +{.} {}							{transition}
    +{}  {\SentenceSpace \PrintReviews}	{review}
}

% The publisher wasn't appearing.
\BibSpec{collection.article}{%
    +{}  {\PrintAuthors}				{author}
    +{,} { \textit}						{title}
    +{.} { }							{part}
    +{:} { \textit}						{subtitle}
    +{,} { \PrintContributions}			{contribution}
    +{,} { \PrintConference}				{conference}
    +{}  {\PrintBook}					{book}
    +{,} { }							{booktitle}
    +{}  { \PrintEditorsB}				{editor}
    +{,} { }							{series}
    +{} { \textbf}						{volume}
    +{,} { }							{publisher}
    +{,} { }							{address}
    +{,} { \PrintDateB}					{date}
    +{,} { }							{pages}
    +{,} { }							{status}
    +{,} { \PrintDOI}					{doi}
    +{,} { available at \eprint}			{eprint}
    +{}  { \parenthesize}				{language}
    +{}  { \PrintTranslation}				{translation}
    +{;} { \PrintReprint}				{reprint}
    +{.} { }							{note}
    +{.} {}							{transition}
    +{}  {\SentenceSpace \PrintReviews}	{review}
}

% Fonts
\usepackage{mathrsfs}
% Enumitem
\usepackage{enumitem}
% TikZ
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
% Hyperrefs
\usepackage{hyperref}

@@

%%%%%%%%%%%%%%%% TIKZ SETTINGS %%%%%%%%%%%%%%%%
\tikzset{every picture/.style={>=stealth},label/.style={font=\footnotesize}}

@@

\begin{abstract}
We consider the rigid monoidal category of character sheaves on a smooth commutative group scheme $G$ over a finite field
$k$ and expand the scope of the function-sheaf dictionary from connected commutative algebraic groups to this setting.
We find the group of isomorphism classes of character sheaves on $G$ and show that it is an extension of the group
of characters of $G(k)$ by a cohomology group determined by the component group scheme of $G$.
We also classify all morphisms in the category character sheaves on $G$.
As an application, we study character sheaves on Greenberg transforms of locally finite type N\'eron models of algebraic tori over local fields. 
This provides a geometrization of quasicharacters of $p$-adic tori.
\end{abstract}

\maketitle

\section*{Introduction}

As Deligne explained in \cite{deligne:SGA4.5}*{Sommes trig.}, if $G$ is a connected commutative algebraic group
over a finite field $k$, then the trace of Frobenius provides a bijection between the group $G(\Fq)^*$ of $\ell$-adic
characters of $G(\Fq)$ and isomorphism classes of those rank-one $\ell$-adic local systems $\mathcal{E}$ on $G$ for which 
\begin{equation}\label{introbox}
m^* \cs{E} \iso \cs{E} \boxtimes \cs{E},
\end{equation}
where $m : G\times G\to G$ is the multiplication map.
If one wishes to make a category from this class of local systems, one is led to consider morphisms
$\cs{E} \to \cs{E}'$ of sheaves which are compatible with particular choices of \eqref{introbox} for $\cs{E}$ and $\cs{E'}$. 
{\it A priori}, the composition $\cs{E} \to \cs{E}' \to \cs{E}''$ of two such morphisms need not be
compatible with the choices of \eqref{introbox} for $\cs{E}$ and $\cs{E}''$.
 However, for connected $G$ the isomorphism \eqref{introbox} is unique, if it exists, and there is no impediment
 to making the dictionary categorical.

If $G$ is a commutative algebraic group over $\Fq$ which is not connected, however, then the isomorphism \eqref{introbox}
need not be unique. In order to track the choice of isomorphism, consider the category $\bCS(G)$ of pairs $(\cs{E},\mu_\cs{E})$
where $\cs{E}$ is a rank-one local system on $G$ and $\mu_\cs{E} : m^*\cs{E} \to \cs{E}\boxtimes\cs{E}$ is a chosen
isomorphism of local systems on $G\times G$. 
In this case, the trace of Frobenius provides an epimorphism from isomorphism classes of objects in $\bCS(G)$ to characters
of $G(\Fq)$, but the epimomorphism need not be injective; consequently,
every character of $G(\Fq)$ may be geometrized as a pair $(\cs{E},\mu_\cs{E})$ but perhaps not uniquely.
Indeed, it follows from a special case of the main result of this paper that the kernel of the trace of Frobenius
$\bCS(G) \to G(\Fq)^*$ trivial if and only if the  group scheme of connected components of $G$ is cyclic.
The defect in the function-sheaf dictionary for characters of commutative algebraic groups over finite fields may be
addressed with the following observation: if $(\cs{E},\mu_\cs{E})$ and $(\cs{E}',\mu_\cs{E}')$ determine the same
character of $G(\Fq)$ then $\cs{E}\iso \cs{E}'$ as local systems on $G$.

Motivated by an application to quasicharacters of algebraic tori over local fields, in this paper we extend the
function-sheaf dictionary from commutative algebraic groups over finite fields to smooth commutative group schemes $G$ over $\Fq$.
In order to do this, we replace the local system $\cs{E}$ on $G$ with a Weil local system while retaining the extra structure $\mu_\cs{E}$. 
In this way we are led to the category $\CS(G)$ of {\it character sheaves} on $G$ (\S\ref{ssec:category}):
objects in $\CS(G)$ are triples $(\gcs{L}, \mu,\phi)$, where $\gcs{L}$ is a rank-one local system on $\bG := G\times_{\Spec{\Fq}} \Spec{\bFq}$
and $\phi : \Frob{G}^* \gcs{L}\to \gcs{L}$ and $\mu : \bm^* \gcs{L} \iso \gcs{L} \boxtimes \gcs{L}$ are
isomorphisms of sheaves satisfying certain compatibility conditions;
morphisms in $\CS(G)$ are then morphisms of Weil sheaves which are compatible with the extra structure.
This paper establishes the basic properties of category $\CS(G)$, using the group homomorphism 
\[
\TrFrob{G}: \CSiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius  to find the relation between character sheaves on $G$ and characters of $G(\Fq)$.
Then we return to our motivating application and use character sheaves to geometrize and categorify quasicharacters algebraic tori over local fields.

We begin our study of category $\CS(G)$ by returning to the case when $G$ is a
connected commutative algebraic group over $\Fq$, revisiting Deligne's function-sheaf dictionary (\S\ref{sec:category}).
We consider character sheaves that arise via base change to $\bFq$ from local systems on $G$ (\S\ref{ssec:descentG}) and
those that appear in a pushforward from a constant sheaf along a discrete isogeny $H \to G$ (\S\ref{ssec:discrete_isogenies}).  
While these constructions make sense even for non-connected $G$, in the connected case we show that
every character sheaf can be described in both of these ways (\S\ref{ssec:connected}).  
We use this fact to prove that $\TrFrob{G} : \CS(G) \to G(\Fq)^*$ is an isomorphism for connected commutative algebraic groups $G$.
We also determine the automorphism groups of character sheaves on such $G$. 
These facts are well known.

Next, we consider character sheaves on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale}).
\'Etale group schemes form a counterpoint to connected algebraic groups, since the component group of any
smooth group scheme is an \'etale group scheme.
Our key tools for understanding the trace of Frobenius in the \'etale case 
are a reinterpretation of $\CS(G)$ in terms of stalks (\S\ref{ssec:stalks}) and 
the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\CSiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT})
that the group homomorphism $\TrFrob{G}: \CSiso{G} \to G(\Fq)^*$ is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
The necessity of using Weil local systems on $G$ in the definition of $\CS(G)$ already appears here:
if one were to use local systems on $G$ instead, the group homomorphism $\TrFrob{G}$ would not then be surjective (\S \ref{ssec:bS}). 
Moreover, as examples show (\S \ref{eg:}), the kernel of $\TrFrob{G}$ is non-trivial in general.

Having understood $\CS(G)$ in two extreme cases --
for connected commutative algebraic groups and for \'etale commutative group schemes --
we turn to the case of smooth commutative group schemes (\S\ref{sec:main}) using the component group sequence
\[
0 \to G^0 \to G \to \pi_0(G) \to 0.
\]
Using pullbacks of character sheaves we obtain a diagram
\[
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar & G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0.
  \end{tikzcd}
\]
We show that the rows of this diagram are exact (\S\S\ref{ssec:restriction},\ref{ssec:component}),
so we may apply the snake lemma to prove the main theorem of the paper,
\begin{theorem*}[{Thm. \ref{thm:snake}}]
If $G$ is a smooth commutative group scheme over $\Fq$ then the trace of Frobenius gives a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
\end{theorem*}
If the component group scheme $\pi_0(\bG)$ is cyclic, then the kernel of $\TrFrob{G}$ will be trivial and
each character of $G(\Fq)$ will uniquely determine a character sheaf on $G$, up to isomorphism.  But when
$\pi_0(\bG)$ is large (c.f. Remark \ref{rem:H2}), $G$ will admit \emph{invisible character sheaves} with
trivial trace of Frobenius.

We also illuminate the nature of the category $\CS(G)$ by showing that every morphism in this category
is either an isomorphism or trivial, and by showing
\begin{theorem*}[{Thm. \ref{thm:autornaught}}]
If $G$ is a smooth commutative group scheme over $\Fq$ then
\[
\Aut(\cs{L}) \iso  \Hh^1(\pi_0(\bG), \EEx)^{\Weil{}}
\]
for all quasicharacter sheaves $\cs{L}$ on $G$.
\end{theorem*}

\subsection*{Application to quasicharacters of \texorpdfstring{$p$}{p}-adic tori and abelian varieties}
As indicated above, our interest in the function-sheaf dictionary for smooth commutative group schemes
over finite fields comes from an application to $p$-adic representation theory,
specifically to quasicharacters (\S\ref{ssec:quasicharacters}) of $p$-adic tori.
However, we found that our method of passing from $p$-adic tori to group schemes over $\Fq$ applies more generally to
any local field $K$ with finite residue field $\Fq$ and to any commutative algebraic group over $K$ that admits a N\'eron model $X$.
This class of algebraic groups over $K$ includes abelian varieties and unipotent $K$-wound groups,
in addition to the algebraic tori we initially considered.

In this paper we show that if $X$ is as above then quasicharacters of $X(K)$ are geometrized and categorified by character sheaves on the
Greenberg transform $\Gr_R(X)$ of the N\'eron model $X$.
Although not locally of finite type, $\Gr_R(X)$ is a commutative group scheme over $\Fq$ and also a projective limit
of smooth commutative group schemes $\Gr^R_n(X)$.
This structure allows us to adapt our work on character sheaves on smooth group schemes over
finite fields to construct (\S\ref{ssec:CS_on_GN}) a category $\QCS(X)$
of {\it quasicharacter sheaves} for $X$, which are certain sheaves on $\Gr_R(X)\times_{\Spec{\Fq}} \Spec{\bFq}$, with extra structure.
The ability to generalize the function-sheaf dictionary to non-connected group schemes plays a crucial role in this application.

Having defined quasicharacter sheaves on N\'eron models of commutative algebraic groups over $K$
and character sheaves on commutative group schemes over $\Fq$, we consider how
these categories are related as $K$ and $\Fq$ vary.  We describe (\S\ref{ssec:basechange}) functors between categories
of quasicharacter sheaves that model restriction and norm homomorphisms of character groups $G(k')^* \to G(k)^*$ and $G(k)^* \to G(k')^*$,
and describe how quasicharacter sheaves behave under Weil restriction (\S\ref{ssec:wrK}).  We also give (\S\ref{ssec:transfer})
a categorical version of a result of Chai and Yu \cite{chai-yu:01a},
relating quasicharacter sheaves for tori over different local fields, even local fields
with different characteristic.

Finally, specializing to the case that $X = T$ is the N\'eron model of an algebraic torus over $K$ (\S\ref{ssec:CS_tori}), 
we give a canonical short exact sequence 
\[
0 \to \Hh^2(X_*(T)_{\mathcal{I}_K},\EEx)^{\Weil{}} \to \QCSiso{T} \to \Hom_\text{}(T(K),\EEx) \to 0,
\]
where $X_*(T)_{\mathcal{I}_K}$ is the group of coinvariants of the cocharacter lattice $X_*(T)$ of the algebraic torus
$T_K$ by the action of the inertia group $\mathcal{I}_K$ of $K$, and where $\Hom_\text{}(T(K),\EEx)$
denotes the group of quasicharacters of $T(K)$.
We further show that automorphism groups in $\QCS(T)$ are given, for every quasicharacter sheaf $\cs{F}$ for $T$, by
\[
\Aut(\cs{F}) \iso (\check{T}_\ell)^{\Weil{K}},
\]
where $\Weil{K}$ is the Weil group for $K$ and $\check{T}_\ell$ is the $\ell$-adic dual torus to $T$.

\subsection*{Relation to other work.}
The main use of the term character sheaf is of course due to Lusztig.
It is applied to certain perverse sheaves on connected reductive algebraic groups over algebraically closed fields in
\cite{lusztig:85a}*{Def.~2.10} and to certain perverse sheaves on certain reductive groups over algebraically closed fields in the series of papers
beginning with \cite{lusztig:disconnected1}.
When commutative, it is not difficult to relate Frobenius-stable character sheaves to our character sheaves
(Remark~\ref{rem:Lusztig}). The new features that we have found pertaining to Weil sheaves
and $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$ do not arise in that context because, for such groups,
Weil sheaves are unnecessary (\S \ref{ssec:alg_groups}) and $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}=0$ (Remark \ref{rem:H2}).

For a connected commutative algebraic group over a finite field, it is not uncommon to refer to
local systems satisfying \eqref{introbox}  as character sheaves; see for example, \cite{kamgarpour:09a}*{Intro}.
Our definition of character sheaves on smooth commutative group schemes over finite fields evolved from this notion,
with an eye toward quasicharacters of $p$-adic groups.
The process of creating a category from the group of quasicharacters of a $p$-adic torus informs our choice of the
term quasicharacter sheaf in this paper.

We anticipate that future work on quasicharacter sheaves will make use of   \cite{suzuki-yoshida:12a} and \cite{suzuki:14a},
and will clarify the relation between this project and other attempts to geometrize admissible distributions on $p$-adic groups,
such as \cite{cunningham-kamgarpour:13a} (limited to quasicharacters of $\ZZ_p^\times$) and \cite{aubert-cunningham:02a}
(limited to characters of depth-zero representations). 
We are actively pursuing the question of how to extend the notion of quasicharacter sheaves to provide a geometrization
of admissible distributions on connected reductive algebraic groups over $p$-adic fields, not just commutative ones.

\subsection*{Acknowledgements.}

We thank Pramod Achar, Masoud Kamgarpour and Hadi Salmasian for 
allowing us to hijack much of a Research in Teams meeting at the Banff International Research Station into a discussion of
quasicharacter sheaves.  We also thank them for their kindness, knowledge and invaluable help.
We thank Takashi Suzuki for clarifying the relation between this project and his recent work and for very helpful
observations and suggestions, especially related to our use of the Hochschild-Serre spectral sequence.
We thank Alessandra Bertrapelle and Cristian Gonz\'ales-Avil\'es for disabusing us of a misapprehension concerning
the Greenberg realization functor and for drawing our attention to their result on Weil restriction and the Greenberg transform.
Finally, we thank Joseph Bernstein for very helpful comments.

Finally, we gratefully acknowledge the financial support of the Pacific Institute for the Mathematical Sciences
and the National Science and Engineering Research Council (Canada), as well the hospitality of the
Banff International Research Station during a Research in Teams program.

\tableofcontents

\section{Definitions and recollections} \label{sec:category}

\subsection{Notations}\label{ssec:notation}

Throughout this paper, $G$ is a smooth commutative group scheme
over a finite field $\Fq$ and $m : G \times G\to G$ is its multiplication morphism.

We will make use of the short exact sequence of smooth group schemes defining the component group scheme for $G$:
\[
\begin{tikzcd}
0 \rar & G^0 \arrow{r}{\iota_0} & G \arrow{r}{\pi_0} & \pi_0(G) \rar & 0.
\end{tikzcd}
\]
Then $G^0$ is a connected algebraic group and $\pi_0(G)$ is an \'etale commutative group scheme.
In contrast to the case of algebraic varieties, the component group scheme $\pi_0(G)$ for $G$ need not be finite.

It follows from the smoothness of $G$ that the structure morphism $G \to \Spec{\Fq}$ is locally of finite type, being smooth.
If the structure morphism $G \to \Spec{\Fq}$ is also \'etale, then $G$ is an \'etale group scheme; this does not imply that $\pi_0(G)$ is finite.
An algebraic group over $\Fq$ is a smooth group scheme of finite type, in which case its component group scheme is finite.

We fix an algebraic closure $\bFq$ of $\Fq$ and write $\bG$ for the
smooth commutative group scheme $G \times_{\Spec{\Fq}} \Spec{\bFq}$ over $\bFq$
obtained by base change from $k$. The multiplication morphism for $\bG$ will be denoted by $\bm$.

Let $\Frob{}$ denote the geometric Frobenius element in $\Gal(\bFq/\Fq)$ as
well as the corresponding automorphism of $\Spec{\bFq}$. The Weil group
$\Weil{}\subset \Gal(\bFq/\Fq)$ is the subgroup generated by $\Frob{}$.
Let $\Frob{G} \ceq \id_{G} \times \Frob{}$ be the Frobenius automorphism of $\bG = G \times_{\Spec{\Fq}} \Spec{\bFq}$.

We fix a prime $\ell$, invertible in $\Fq$.
We will work with constructible $\ell$-adic sheaves \citelist{\cite{deligne:80a}*{\S 1.1} \cite{SGA5}*{Expos\'es V, VI}}
on schemes locally of finite type over $\Fq$, employing the standard formalism.
We also make extensive use of the external tensor product of $\ell$-adic sheaves,
defined as follows: if $\mathcal{F}$ and $\mathcal{G}$ are constructible $\ell$-adic
sheaves on schemes $X$ and $Y$ and $p_X : X\times Y\to X$ and $p_Y : X\times Y \to Y$
are the projections, then $\mathcal{F}\boxtimes \mathcal{G} \ceq p_X^* \mathcal{F} \otimes p_Y^*\mathcal{G}$.

For any commutative group $A$, we will write $A^*$ for the dual group $\Hom(A, \EEx)$.

\subsection{Character sheaves on commutative group schemes over finite fields}\label{ssec:category}

\begin{definition}\label{def:CS}
A \emph{character sheaf on $G$} is a triple
$\cs{L}\ceq (\gcs{L},\mu,\phi)$ where:
\begin{enumerate}
\labitem{(CS.1)}{CS.1} $\gcs{L}$ is a rank-one $\ell$-adic local system on $\bG$, by which we mean a constructible
$\ell$-adic sheaf on $\bG$, {\it lisse} on each connected component of $\bG$, whose stalks are one-dimensional $\EE$-vector spaces;
\labitem{(CS.2)}{CS.2} $\mu: \bm^* \gcs{L} \to \LxL$ is an isomorphism of
sheaves on $\bGxG$ such that the following diagram commutes,
  where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;
  \[
  \begin{tikzcd}[row sep=30]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL
  \end{tikzcd}
  \]
\labitem{(CS.3)}{CS.3} $\phi : \Frob{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of constructible $\ell$-adic sheaves on $\bG$ compatible with
  $\mu$ in the sense that the following diagram commutes.
  \[
  \begin{tikzcd}[row sep=20]
  \Frob{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob{\GxxG}^*\mu}
    && \Frob{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob{G}^* \gcs{L} \arrow[swap]{d}{\bm^* \phi}
    && \Frob{G}^*\gcs{L}\boxtimes \Frob{G}^*\gcs{L} \dar{\phi\boxtimes \phi} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL
  \end{tikzcd}
  \]
\end{enumerate}
Morphisms of character sheaves are defined in the natural way:
\begin{enumerate}
\labitem{(CS.4)}{CS.4} if $\cs{L} = (\gcs{L},\mu,\phi)$ and
  $\cs{L'} = (\gcs{L'},\mu',\phi')$ are character sheaves on $G$ then
  a morphism $\rho : \cs{L} \to \cs{L}'$ is a map $\brho : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that the following diagrams both commute.
  \[
  \begin{tikzcd}[column sep=40]
  \Frob{G}^* \gcs{L} \rar{\Frob{G}^* \brho} \arrow[swap]{d}{\phi} & \Frob{G}^* \gcs{L'} \dar{\phi'}
  & & \arrow[swap]{d}{\mu} \bm^* \gcs{L} \rar{\bm^* \brho} & \bm^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\brho} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\rho\boxtimes \rho}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
\end{enumerate}
 The category of character sheaves on $G$ will be denoted by $\CS(G)$.
 \end{definition}

Category $\CS(G)$ is a rigid monoidal category
\cite{etingof:09a}*{\S1.10} under the tensor product
$\cs{L} \otimes \cs{L'}$ defined by $(\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', \phi\otimes \phi')$,
with duals given by applying the sheaf hom functor
$\sheafHom(\ - \ ,\EE)$.
This rigid monoidal category structure for $\CS(G)$ gives the set $\CSiso{G}$
of isomorphism classes in $\CS(G)$ the structure of a group.

\begin{remark}
The category of character sheaves on $G$ is not abelian since it is not closed under direct sums; 
thus $\CS(G)$ is not a tensor category in the sense of \cite{deligne:02a}*{0.1}.  
We suspect that requiring that $\mu$ be injective rather than
an isomorphism and dropping the condition that the stalks be one-dimensional would yield an abelian category.
\end{remark}

We will describe the group $\CSiso{G}$ in Theorem~\ref{thm:snake}
and the sets $\Hom(\cs{L},\cs{L}')$ in Theorem~\ref{thm:autornaught}; in this way we provide a complete description of the category $\CS(G)$.
In the meantime, we make an elementary observation about $\Hom(\cs{L},\cs{L}')$.

\begin{lemma}\label{lem:autornaught}
Let $G$ be a smooth commutative group scheme over $\Fq$.
If $\cs{L}$ and $\cs{L}'$ are character sheaves on $G$, then
every $\rho\in \Hom(\cs{L},\cs{L}')$ is either trivial (zero on every stalk)
or an isomorphism. 
\end{lemma}

\iffalse
\begin{proof}
Suppose $\rho \in \Hom(\cs{L},\cs{L}')$.
We prove the lemma by considering the linear transformations $\brho_{\bg} : \gcs{L}_{\bg} \to \gcs{L}_{\bg}$ at the stalks
above geometric points $\bg$ on $G$ and showing that, either each $\brho_{\bg}$ trivial or each $\brho_{\bg}$ is an isomorphism.
(This idea is expanded upon in Section~\ref{ssec:stalks}.)
Let ${\bar e}$ be the geometric point above the identity $e$ for $G$ determined by our choice of algebraic closure $\bFq$ of $\Fq$.
If $\brho_{\bar e} = 0$ then the second diagram in \ref{CS.4} implies that $\brho_{\bg} = 0$ for all $\bg$, in which case $\rho$ is trivial.
On the other hand, if $\brho_{\bar e}$ is non-trivial then the second diagram in \ref{CS.4} implies that $\brho_{\bg}$ is non-trivial
for all $\bg$ and thus an isomorphism, since the stalks of character sheaves are one-dimensional; in this case $\rho$ is an isomorphism.
\end{proof}
\fi

\subsection{Trace of Frobenius}\label{ssec:Frob}

In this section we introduce two tools which will help us understand isomorphism classes of objects in $\CS(G)$:
the map $\CSiso{G} \to G(k)^*$ given by trace of Frobenius and the pullback functor $\CS(G) \to \CS(H)$
associated to a morphism $H \to G$ of smooth group schemes over $\Fq$.

Let $(\gcs{L},\phi)$ be a Weil sheaf on $G$. Every $g\in G(\Fq)$
determines a geometric point $\bg$ fixed by $\Frob{G}$. 
Together with the canonical isomorphism $(\Frob{G}^*\gcs{L})_{\bg} \iso  \gcs{L}_{\Frob{G}(\bg)}$,
the automorphism $\phi$ determines an automorphism $\phi_{\bg}$ of the $\EE$-vector space $\gcs{L}_{\bg}$.
Let $\trace(\phi_{\bg};\gcs{L}_{\bg})$ be the trace of $\phi_{\bg} \in \Aut_{\EE}(\gcs{L}_{\bg})$ and let
$\trFrob{(\gcs{L},\phi)} : G(\Fq)\to \EE$ be the function defined by 
\begin{equation}\label{trWeil}
\trFrob{(\gcs{L},\phi)}(g) \ceq \trace(\phi_{\bg};\gcs{L}_{\bg}),
\end{equation}
commonly called the {\em trace of Frobenius of $(\gcs{L},\phi)$}.
Note that if $(\gcs{L},\phi) \iso (\gcs{L'},\phi')$ as Weil sheaves, 
then $\trFrob{(\gcs{L},\phi)} = \trFrob{(\gcs{L'},\phi')}$ as functions on $G(\Fq)$.

Now suppose $\cs{L} = (\gcs{L},\mu,\phi)$ is a character sheaf on $G$.
Then the isomorphism $\bm^* \gcs{L} \iso \gcs{L} \boxtimes\gcs{L}$ and the diagram of
\ref{CS.3} guarantee
that the function $\trFrob{(\gcs{L},\phi)} : G(\Fq)\to \EEx$ is a group homomorphism, which we will also denote by $\trFrob{\cs{L}}$.
Moreover, this homomorphism depends only on the isomorphism class of $\cs{L}$, so we obtain a map
\begin{align*}
\TrFrob{G} : \CSiso{G} &\to G(\Fq)^*, \\
\cs{L} &\mapsto \trFrob{\cs{L}}.
\end{align*}
Since tensor products on the stalks of $\cs{L}$ induce pointwise multiplication on the trace of Frobenius, $\TrFrob{G}$ is a group homomorphism.

The next two results follow easily from the definitions.

\begin{lemma}\label{lem:pullback}
  If $f : H\to G$ is a morphism of smooth commutative group schemes over $\Fq$, then
  \begin{align*}
  f^* : \CS(G) &\to \CS(H) \\
  (\gcs{L},\mu,\phi) &\mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)
  \end{align*}
  defines a monoidal functor dual to $f \colon H(\Fq) \to G(\Fq)$ in the sense that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{\TrFrob{G}} & \CSiso{H} \dar{\TrFrob{H}} \\
   G(\Fq)^* \rar & H(\Fq)^*
  \end{tikzcd}
  \]
  is a commutative diagram of groups.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{lemma}
\iffalse
\begin{proof}
  Let $\cs{L}$ be a character sheaf on $G$. 
  Pullback by $\bf$ takes rank-one local systems to rank-one local systems.
  To see that $(\bfxf)^* \mu$ satisfies \ref{CS.2},
  apply the functor $(\bfxf)^*$
  to \ref{CS.2} for $\cs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gcs{L}$.
  To show that $f^*\cs{L}$ satisfies
  \ref{CS.3}, apply the same functor to \ref{CS.3} for $\cs{L}$.
  Since $f$ is a morphism of group schemes defined over $\Fq$
  it provides isomorphisms $(\bfxf)^*\Frob{\GxxG}^* \iso \Frob{\GxxG}^* (\bfxf)^*$
  and $(\bfxf)^* \bm^*\iso \bm^* \bf^*$ between functors of constructible sheaves.

  Applying $\bf^*$ and $\bf^* \tight{1}{\times}{1}\bf^*$ to \ref{CS.4} defines the action
  of $f^*$ on morphisms of character sheaves; arguing as above shows that $f^*$ is
  a functor from $\CS(G)$ to $\CS(H)$.  Since tensor products commute with pullback in schemes,
  $f^* : \CS(G) \to \CS(H)$ is a monoidal functor.
  The diagram relating $f^* : \CS(G) \to \CS(H)$, $f^* : G(k)^* \to H(k)^*$ and trace of Frobenius
  commutes by \cite{laumon:87a}*{1.1.1.2}, where the ambient
 finite type hypothesis can be replaced by locally of finite type.

  Finally, the fact that $(f\circ g)^* = g^* \circ f^*$ follows from the analogous
  statements about the pullback functor on $\ell$-adic constructible sheaves.
\end{proof}
\fi

If $G_1$ and $G_2$ are smooth commutative group schemes over $\Fq$ then characters of $(G_1 \times G_2)(\Fq)$
all take the form $\chi_1\otimes \chi_2$ for characters $\chi_1$ of $G_1(\Fq)$ and $\chi_2$ of $G_2(\Fq)$. 
The next lemma shows that character sheaves on $G$ enjoy an analogous property.

>>>>

\documentclass[10pt]{amsart}

\title{Commutative Character Sheaves}
\author{Clifton Cunningham}
\address{Department of Mathematics and Statistics, University of Calgary, 2500 University Drive Northwest, Calgary, AB, Canada, {T2N~1N4}.}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Department of Mathematics, University of Pittsburgh, 301 Thackeray Hall, Pittsburgh, PA , United States, 15260.}
\email{roed.math@gmail.com}

\subjclass[2010]{14F05 (primary), 14L15 (secondary), 22E50 (tertiary)}
\keywords{character sheaves}

\usepackage{amssymb}
\usepackage{amsrefs}
% Fonts
\usepackage{mathrsfs}
% Enumitem
\usepackage{enumitem}
% Hyperrefs
\usepackage{hyperref}

@@

@@

\section{Commutative Character Sheaves}

\section{No Invisible Character Sheaves}

By the universal coefficient theorem

In the three examples above, we are able to compute $\Hh^0(\Weil{}, \Hh^2(\bG, \EEx))$ without reference to $\Frob{G}$ as a consequence of the structure of $\Hh^0(\Weil{}, \Hh^2(\bG, \EEx))$; this independence does not hold in general.  But the description of $\Hh^2(\bG, \EEx)$ allows us to
